---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
set.seed(1)
```

# psychtm: A package for text mining in psychological research

<!-- badges: start -->
[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
[![R-CMD-check](https://github.com/ktw5691/psychtm/workflows/R-CMD-check/badge.svg)](https://github.com/ktw5691/psychtm/actions)
![CRAN status](https://www.r-pkg.org/badges/version/psychtm)
[![Codecov test coverage](https://codecov.io/gh/ktw5691/psychtm/branch/main/graph/badge.svg)](https://app.codecov.io/gh/ktw5691/psychtm?branch=main)
<!-- badges: end -->

<!-- [![CRAN status](https://www.r-pkg.org/badges/version/psychtm)](https://CRAN.R-project.org/package=psychtm) -->

The goal of `psychtm` is to make text mining models and methods accessible for social science researchers, particularly within psychology. This package allows users to

-   Estimate the SLDAX topic model and popular models subsumed by SLDAX, including SLDA, LDA, and regression models;

-   Obtain posterior inferences;

-   Assess model fit using coherence and exclusivity metrics.

## Installation

Once on CRAN, install the package as usual:

``` r
install.packages("psychtm")
```

Alternatively, you can install the most current development version:

-   If necessary, first install the `devtools` R package,

``` r
install.packages("devtools")
```

### Option 1: Install the latest stable version from Github

``` r
devtools::install_github("ktw5691/psychtm")
```

### Option 2: Install the latest development snapshot

``` r
devtools::install_github("ktw5691/psychtm@devel")
```

## Example

This is a basic example which shows you how to (1) prepare text documents stored in a data frame; (2) fit a supervised topic model with covariates (SLDAX); and (3) summarize the regression relationships from the estimated SLDAX model.

```{r example}
library(psychtm)
library(lda) # Required if using `prep_docs()`

data(teacher_rate)  # Synthetic student ratings of instructors
docs_vocab <- prep_docs(teacher_rate, "doc")
vocab_len <- length(docs_vocab$vocab)
fit_sldax <- gibbs_sldax(rating ~ I(grade - 1),
                         data = teacher_rate,
                         docs = docs_vocab$documents,
                         V = vocab_len,
                         K = 2,
                         model = "sldax")
eta_post <- post_regression(fit_sldax)
```

```{r reg_summary}
summary(eta_post)
```

For a more detailed example of the key functionality of this package, explore the vignette(s) for a good starting point:

``` r
browseVignettes("psychtm")
```

## How to Cite the Package

Wilcox, K. T., Jacobucci, R., Zhang, Z., Ammerman, B. A. (2021). Supervised latent Dirichlet allocation with covariates: A Bayesian structural and measurement model of text and covariates. *PsyArXiv*. <https://doi.org/10.31234/osf.io/62tc3>

## Common Troubleshooting

Ensure that appropriate `C++` compilers are installed on your computer:

-   Mac users will have to download [Xcode](https://apps.apple.com/ca/app/xcode/id497799835?mt=12) and its related Command Line Tools (found within Xcode's Preference Pane under Downloads/Components).

-   Windows users may need to install [Rtools](https://CRAN.R-project.org/bin/windows/Rtools/). For easier command line use, be sure to select the option to install Rtools to their path.

-   Most Linux distributions should already have up-to-date compilers.

## Limitations

-   This package uses a Gibbs sampling algorithm that can be memory-intensive for a large corpus.

## Getting Help

If you think you have found a bug, please [open an  issue](https://github.com/ktw5691/psychtm/issues) and provide a [minimal complete verifiable example](https://stackoverflow.com/help/mcve).
